### 2021-12-13 draft_yma.Rmd
marginal plot

---
title: "draft_yma"
author: "Anna Ma"
date: "12/12/2021"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readr)
```

## Data exploration

 **converted variables to per person and docs, beds, etc.**

```{r, message=FALSE}
cdi = read_csv("cdi.csv") %>%
  janitor::clean_names() %>%
  mutate(crm_1000 = 1000*(crimes/pop),
         pdocs_1000 = 1000*(docs/pop),
         pbeds_1000 = 1000*(beds/pop),
         density_pop = pop/area,
         # convert region to factors and recoded them accordingly 
         region = factor(region, levels = 1:4,
                    labels = c("northeast", "north central", "south", "west"))) %>% select(-c(docs,beds)) %>% view()

```

### Descriptive statistics of all varaibles
```{r statistic summary}

cdi_descriptive = cdi %>% select(-c(id,cty,state,region))

# Global
skimr::skim(cdi_descriptive) %>% 
  select(-c("skim_type","complete_rate")) %>% 
    mutate(skim_variable = 
             recode(skim_variable, pcincome = "pcincome (in dollars)", totalinc = "totalinc (in million of dollars)"
  )) %>% 
  knitr::kable(
    col.names = c("variable", "n_missing", "mean","sd","min","Q25","median","Q75","max","histogram"),
    caption = "Global Summary", digits = 4)
```

Q: 1. do we need to group them by state or county? group by gounty gives 5000+ rows though...
  2. do we need box plot still? 
  3. again, transfer some variables to "per pop" / "per 1000 pop" ?

### Descriptive analysis

We can use the box plot/ or histogram to check for normality. But I forgot when do we need normality...isn't it for residual??

```{r}
par(mfrow = c(3,5))

boxplot(cdi$density_pop,  main = "density_pop")
boxplot(cdi$area, main = "area")

boxplot(cdi$pop, main = "pop")
boxplot(cdi$pop18, main = "pop18")
boxplot(cdi$pop65, main = "pop65")

boxplot(cdi$pdocs_1000, main = "pdocs_1000")
boxplot(cdi$pbeds_1000, main = "pbeds_1000")

boxplot(cdi$crm_1000,  main = "crm_1000")

boxplot(cdi$hsgrad, main = "hsgrad")
boxplot(cdi$bagrad, main = "bagrad")
boxplot(cdi$poverty, main = "poverty")
boxplot(cdi$unemp,main = "unemp")
boxplot(cdi$pcincome, main = "pcincome")
boxplot(cdi$totalinc, main = "totalinc")

```

### Correlation

#### Pairwise relationship

* This gives us an idea of the correlation between each variable, but my old project build whole model first then assessed the correlation. Need Discussion!

```{r}
library(corrplot)
corrplot(cor(cdi_descriptive), type = "upper", diag = FALSE)
```

* Here is the correlation with pairs function. More specific than the heat map above. We can observe the correlation between all the terms here
```{r}
pairs(crm_1000 ~.,data=cdi_descriptive, panel = panel.smooth, upper.panel = NULL, main = "Scatterplot Matrix")
```

#### Marginal distribution ?

```{r}
library(ggplot2)
library(ggExtra)

```

```{r}
marg_den = cdi %>% ggplot(aes(x = density_pop, y = crm_1000)) + geom_point(alpha = 0.3) + geom_smooth(method = 'lm', se = TRUE, color = 'red')
ggMarginal(marg_den, type = "histogram", fill="transparent")
```

```{r}
marg_area = cdi %>% ggplot(aes(x = area, y = crm_1000)) + geom_point(alpha = 0.3) + geom_smooth(method = 'lm', se = TRUE, color = 'red')
ggMarginal(marg_area, type = "histogram", fill="transparent")
```

```{r}
marg_pop = cdi %>% ggplot(aes(x = pop, y = crm_1000)) + geom_point(alpha = 0.3) + geom_smooth(method = 'lm', se = TRUE, color = 'red')
ggMarginal(marg_pop, type = "histogram", fill="transparent")
```

```{r}
marg_pop18 = cdi %>% ggplot(aes(x = pop18, y = crm_1000)) + geom_point(alpha = 0.3) + geom_smooth(method = 'lm', se = TRUE, color = 'red')
# positive correlation
ggMarginal(marg_pop18, type = "histogram", fill="transparent")
```

```{r}
marg_pop65 = cdi %>% ggplot(aes(x = pop65, y = crm_1000)) + geom_point(alpha = 0.3) + geom_smooth(method = 'lm', se = TRUE, color = 'red')
ggMarginal(marg_pop65, type = "histogram", fill="transparent")
```

```{r}
marg_pdocs_1000 = cdi %>% ggplot(aes(x = pdocs_1000, y = crm_1000)) + geom_point(alpha = 0.3) + geom_smooth(method = 'lm', se = TRUE, color = 'red')
ggMarginal(marg_pdocs_1000, type = "histogram", fill="transparent")
```

```{r}
marg_pbeds_1000 = cdi %>% ggplot(aes(x = pbeds_1000, y = crm_1000)) + geom_point(alpha = 0.3) + geom_smooth(method = 'lm', se = TRUE, color = 'red')
ggMarginal(marg_pbeds_1000, type = "histogram", fill="transparent")
```

```{r}
marg_hsgrad = cdi %>% ggplot(aes(x = hsgrad, y = crm_1000)) + geom_point(alpha = 0.3) + geom_smooth(method = 'lm', se = TRUE, color = 'red') #negative correlation
ggMarginal(marg_hsgrad, type = "histogram", fill="transparent")
```

```{r}
marg_bagrad = cdi %>% ggplot(aes(x = bagrad, y = crm_1000)) + geom_point(alpha = 0.3) + geom_smooth(method = 'lm', se = TRUE, color = 'red')
ggMarginal(marg_bagrad, type = "histogram", fill="transparent")
```

```{r}
marg_poverty = cdi %>% ggplot(aes(x = poverty, y = crm_1000)) + geom_point(alpha = 0.3) + geom_smooth(method = 'lm', se = TRUE, color = 'red') # positive correlation
ggMarginal(marg_poverty, type = "histogram", fill="transparent")
```

```{r}
marg_unemp = cdi %>% ggplot(aes(x = unemp, y = crm_1000)) + geom_point(alpha = 0.3) + geom_smooth(method = 'lm', se = TRUE, color = 'red')
ggMarginal(marg_unemp, type = "histogram", fill="transparent")
```

```{r}
marg_pcincome = cdi %>% ggplot(aes(x = pcincome, y = crm_1000)) + geom_point(alpha = 0.3) + geom_smooth(method = 'lm', se = TRUE, color = 'red')
ggMarginal(marg_pcincome, type = "histogram", fill="transparent")
```

```{r}
marg_totalinc = cdi %>% ggplot(aes(x = totalinc, y = crm_1000)) + geom_point(alpha = 0.3) + geom_smooth(method = 'lm', se = TRUE, color = 'red')
ggMarginal(marg_totalinc, type = "histogram", fill="transparent")
```

### Distribution of outcome

```{r}
cdi %>% 
  ggplot(aes(x = crm_1000)) +
  geom_histogram()

# log transfor the outcome
cdi %>% 
  ggplot(aes(x = log(crm_1000))) +
  geom_histogram()

```

do we look at the distribution of outcome like this and transform them here? check again

### States with unusual rates

group by state
```{r}
cdi_by_state = cdi %>% group_by(state) %>% 
  summarise(state_pop = sum(pop),
            state_crimes = sum(crimes)) %>% 
  mutate(state_CRM_1000 = 1000*state_crimes/state_pop)
```

Not sure what counts as out lier, this is what the other group did.

```{r}
upper = quantile(cdi$crm_1000, 0.75)
lower = quantile(cdi$crm_1000, 0.25)
IQR = upper - lower
cdi %>% 
  filter(crm_1000 > upper + 1.5*IQR,
         crm_1000 > lower - 1.5*IQR) %>% 
  dplyr::select(cty, crm_1000) %>%
  knitr::kable(digits = 2)
```

This shows no outlier on the state level tho...

```{r}
upper = quantile(cdi_by_state$state_CRM_1000, 0.75)
lower = quantile(cdi_by_state$state_CRM_1000, 0.25)
IQR = upper - lower
cdi_by_state %>% 
  filter(state_CRM_1000 > upper + 1.5*IQR,
         state_CRM_1000 > lower - 1.5*IQR) %>% 
  dplyr::select(state, state_CRM_1000) %>%
  knitr::kable(digits = 2)
boxplot(cdi_by_state$state_CRM_1000, main = 'State Crime Rate per 1000 people')
```