### 2021-12-16 ref.Rmd

---
title: "Final Project"
author: "Jiacheng Wu"
date: "12/2/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(corrplot)
library(leaps)
library(MASS)
library(performance)
```

```{r}
crime_df = read.csv("cdi.csv")
```

```{r}
crime_df_1 = 
  dplyr::select(crime_df, -id, -cty, -state)

crime_df_1 = 
  crime_df_1 %>% 
  mutate(CRM.1000 = crimes*1000/pop) %>% 
  dplyr::select(-crimes)
```

```{r}
pairs(crime_df_1)
```

```{r}
corrplot(cor(crime_df_1), type = "upper", diag = FALSE)
```

```{r}
boxplot(crime_df_1$pop, main = "Total population")
boxplot(crime_df_1$pop18, main = "Percent of population aged 18-34")
boxplot(crime_df_1$pop65, main = "Percent of population aged 65+")
boxplot(crime_df_1$area, main = "Land area")
boxplot(crime_df_1$docs, main = "Number of active physicians")
boxplot(crime_df_1$beds, main = "Number of hospital beds")
boxplot(crime_df_1$CRM.1000, main = "Total serious crimes")
boxplot(crime_df_1$hsgrad, main = "Percent high school graduates")
boxplot(crime_df_1$bagrad, main = "Percent bachelor's degrees")
boxplot(crime_df_1$poverty, main = "Percent below poverty level")
boxplot(crime_df_1$unemp, main = "Percent unemployment")
boxplot(crime_df_1$pcincome, main = "Per capita income")
boxplot(crime_df_1$totalinc, main = "Total personal income")
boxplot(crime_df_1$region, main = "Geographical region")

mult.fit1 = lm(CRM.1000 ~ ., data = crime_df_1)
summary(mult.fit1)
```

## Model diagnostics

## Residual vs Fitted & QQ Plots
```{r}

# residual vs fitted plot
plot(mult.fit1, which = 1)

# QQ plot
plot(mult.fit1, which = 2)

# read new data with Ln transformation
crime_df = read.csv("cdi.csv")
crime_df_2 = 
  dplyr::select(crime_df, -id, -cty, -state)

# fit model - log transform the outcome
mult.fit2 = lm(log(CRM.1000) ~ ., data = crime_df_1)

# residual vs fitted plot
plot(mult.fit2, which = 1)

# QQ plot
plot(mult.fit2, which = 2)

```

## Backwards Elimination

Take out non-significant variables one at a time starting with the highest p-value.

```{r}

step1 = update(mult.fit1, . ~ . -hsgrad)
summary(step1)

step2 = update(step1, . ~ . -unemp)
summary(step2)

step3 = update(step2, . ~ . -bagrad)
summary(step3)

step4 = update(step3, . ~ . -docs)
summary(step4)

step5 = update(step4, . ~ . -pop65)
summary(step5)

step(mult.fit1, direction='backward')
step
```

## Forward Selection

Reversed backward elimination starting with the lowest p-value.
### Step 1:  Fit simple linear regressions for all variables,look for the variable with lowest p-value
```{r}

fit1 = lm(CRM.1000 ~ area, data = crime_df_1)
summary(fit1)
fit2 = lm(CRM.1000 ~ pop, data = crime_df_1)
summary(fit2)
fit3 = lm(CRM.1000 ~ pop18, data = crime_df_1)
summary(fit3)
fit4 = lm(CRM.1000 ~ pop65, data = crime_df_1)
summary(fit4)
fit5 = lm(CRM.1000 ~ docs, data = crime_df_1)
summary(fit5)
fit6 = lm(CRM.1000 ~ beds, data = crime_df_1)
summary(fit6)
fit7 = lm(CRM.1000 ~ hsgrad, data = crime_df_1)
summary(fit7)
fit8 = lm(CRM.1000 ~ bagrad, data = crime_df_1)
summary(fit8)
fit9 = lm(CRM.1000 ~ poverty, data = crime_df_1)
summary(fit9)
fit10 = lm(CRM.1000 ~ unemp, data = crime_df_1)
summary(fit10)
fit11 = lm(CRM.1000 ~ pcincome, data = crime_df_1)
summary(fit11)
fit12 = lm(CRM.1000 ~ totalinc, data = crime_df_1)
summary(fit12)
fit13 = lm(CRM.1000 ~ region, data = crime_df_1)
summary(fit13)
```
# Enter first the one with the lowest p-value: poverty
```{r}
forward1 = lm(CRM.1000 ~ poverty, data = crime_df_1)
summary(forward1)

### Step 2: Enter the one with the lowest p-value in the rest 
fit1 = update(forward1, . ~ . +area)
summary(fit1)
fit2 = update(forward1, . ~ . +pop)
summary(fit2)
fit3 = update(forward1, . ~ . +pop18)
summary(fit3)
fit4 = update(forward1, . ~ . +pop65)
summary(fit4)
fit5 = update(forward1, . ~ . +docs)
summary(fit5)
fit6 = update(forward1, . ~ . +beds)
summary(fit6)
fit7 = update(forward1, . ~ . +hsgrad)
summary(fit7)
fit8 = update(forward1, . ~ . +bagrad)
summary(fit8)
fit9 = update(forward1, . ~ . +unemp)
summary(fit9)
fit10 = update(forward1, . ~ . +pcincome)
summary(fit10)
fit11 = update(forward1, . ~ . +totalinc)
summary(fit11)
fit12 = update(forward1, . ~ . +region)
summary(fit12)
```
# Enter the one with the lowest p-value: beds
```{r}
forward2 = update(forward1, . ~ . + beds)
summary(forward2)
### Step 3: Enter the one with the lowest p-value in the rest 
fit1 = update(forward2, . ~ . +area)
summary(fit1)
fit2 = update(forward2, . ~ . +pop)
summary(fit2)
fit3 = update(forward2, . ~ . +pop18)
summary(fit3)
fit4 = update(forward2, . ~ . +pop65)
summary(fit4)
fit5 = update(forward2, . ~ . +docs)
summary(fit5)
fit6 = update(forward2, . ~ . +hsgrad)
summary(fit6)
fit7 = update(forward2, . ~ . +bagrad)
summary(fit7)
fit8 = update(forward2, . ~ . +unemp)
summary(fit8)
fit9 = update(forward2, . ~ . +pcincome)
summary(fit9)
fit10 = update(forward2, . ~ . +totalinc)
summary(fit10)
fit11 = update(forward2, . ~ . +region)
summary(fit11)
```

# Enter the one with the lowest p-value: region

```{r}
forward3 = update(forward2, . ~ . + region)
summary(forward3)
### Step 4: Enter the one with the lowest p-value in the rest 
fit1 = update(forward3, . ~ . +area)
summary(fit1)
fit2 = update(forward3, . ~ . +pop)
summary(fit2)
fit3 = update(forward3, . ~ . +pop18)
summary(fit3)
fit4 = update(forward3, . ~ . +pop65)
summary(fit4)
fit5 = update(forward3, . ~ . +docs)
summary(fit5)
fit6 = update(forward3, . ~ . +hsgrad)
summary(fit6)
fit7 = update(forward3, . ~ . +bagrad)
summary(fit7)
fit8 = update(forward3, . ~ . +unemp)
summary(fit8)
fit9 = update(forward3, . ~ . +pcincome)
summary(fit9)
fit10 = update(forward3, . ~ . +totalinc)
summary(fit10)
```
# Enter the one with the lowest p-value: totalinc
```{r}
forward4 = update(forward3, . ~ . + totalinc)
summary(forward4)
### Step 4: Enter the one with the lowest p-value in the rest 
fit1 = update(forward4, . ~ . +area)
summary(fit1)
fit2 = update(forward4, . ~ . +pop)
summary(fit2)
fit3 = update(forward4, . ~ . +pop18)
summary(fit3)
fit4 = update(forward4, . ~ . +pop65)
summary(fit4)
fit5 = update(forward4, . ~ . +docs)
summary(fit5)
fit6 = update(forward4, . ~ . +hsgrad)
summary(fit6)
fit7 = update(forward4, . ~ . +bagrad)
summary(fit7)
fit8 = update(forward4, . ~ . +unemp)
summary(fit8)
fit9 = update(forward4, . ~ . +pcincome)
summary(fit9)
```
# Enter the one with the lowest p-value: pcincome
```{r}
forward5 = update(forward4, . ~ . + pcincome)
summary(forward4)
### Step 5: Enter the one with the lowest p-value in the rest 
fit1 = update(forward5, . ~ . +area)
summary(fit1)
fit2 = update(forward5, . ~ . +pop)
summary(fit2)
fit3 = update(forward5, . ~ . +pop18)
summary(fit3)
fit4 = update(forward5, . ~ . +pop65)
summary(fit4)
fit5 = update(forward5, . ~ . +docs)
summary(fit5)
fit6 = update(forward5, . ~ . +hsgrad)
summary(fit6)
fit7 = update(forward5, . ~ . +bagrad)
summary(fit7)
fit8 = update(forward5, . ~ . +unemp)
summary(fit8)
```
# Enter the one with the lowest p-value: pop
```{r}
forward6 = update(forward5, . ~ . + pop)
summary(forward5)
### Step 6: Enter the one with the lowest p-value in the rest 
fit1 = update(forward6, . ~ . +area)
summary(fit1)
fit2 = update(forward6, . ~ . +pop18)
summary(fit2)
fit3 = update(forward6, . ~ . +pop65)
summary(fit3)
fit4 = update(forward6, . ~ . +docs)
summary(fit4)
fit5 = update(forward6, . ~ . +hsgrad)
summary(fit5)
fit6 = update(forward6, . ~ . +bagrad)
summary(fit6)
fit7 = update(forward6, . ~ . +unemp)
summary(fit7)
```

# Enter the one with the lowest p-value: pop18

```{r}
forward7 = update(forward6, . ~ . + pop18)
summary(forward7)
### Step 6: Enter the one with the lowest p-value in the rest 
fit1 = update(forward7, . ~ . +area)
summary(fit1)
fit2 = update(forward7, . ~ . +pop65)
summary(fit2)
fit3 = update(forward7, . ~ . +docs)
summary(fit3)
fit4 = update(forward7, . ~ . +hsgrad)
summary(fit4)
fit5 = update(forward7, . ~ . +bagrad)
summary(fit5)
fit6 = update(forward7, . ~ . +unemp)
summary(fit6)
```

# Enter the one with the lowest p-value:area 
```{r}
forward8 = update(forward7, . ~ . + area)
summary(forward8)
### Step 8: Enter the one with the lowest p-value in the rest 
fit1 = update(forward8, . ~ . +pop65)
summary(fit1)
fit2 = update(forward8, . ~ . +docs)
summary(fit2)
fit3 = update(forward8, . ~ . +hsgrad)
summary(fit3)
fit4 = update(forward8, . ~ . +bagrad)
summary(fit4)
fit5 = update(forward8, . ~ . +unemp)
summary(fit5)
```

# P-value of all new added variables are larger than 0.05, which means that they 
# are not significant predictor, and we stop here.

```{r}
# The model we obtained is CRM.1000 ~ poverty + beds + region + totalinc + pcincome + pop + pop18 + area
mult.fit1.final = lm(CRM.1000 ~ poverty + beds + region + totalinc + pcincome + pop + pop18 + area, data = crime_df_1)
summary(mult.fit1.final)

# fit using one function
step(mult.fit1, direction = 'forward')
```

## Stepwise

```{r}
step.model = stepAIC(mult.fit1, direction = "both", 
                      trace = FALSE)
summary(step.model)
```

## Criterion based

```{r}
mat = as.matrix(crime_df_1)

leaps(x = mat[,1:13], y = mat[,14], nbest = 1, method = "Cp")

leaps(x = mat[,1:13], y = mat[,14], nbest = 1, method = "adjr2")

# Function regsubsets() performs a subset selection by identifying the "best" model that contains
# a certain number of predictors. By default "best" is chosen using SSE/RSS (smaller is better)
b = regsubsets(CRM.1000 ~ ., data = crime_df_1)
rs = summary(b)

# plot of Cp and Adj-R2 as functions of parameters
par(mfrow=c(1,2))

plot(1:8, rs$cp, xlab="No of parameters", ylab="Cp Statistic")
abline(0,1)

plot(1:8, rs$adjr2, xlab="No of parameters", ylab="Adj R2")
```

## Diagnosing stepwise model

```{r}
plot(step.model, which = 1)
plot(step.model, which = 2)

# fit model - log transform the outcome
step.modellog = lm(log(CRM.1000) ~ ., data = crime_df_1)

# residual vs fitted plot
plot(step.modellog, which = 1)

# QQ plot
plot(step.modellog, which = 2)

```

```{r}
boxcox(step.model, lambda = seq(-3, 3, by = 0.25))
step.modelsqrt = lm(sqrt(CRM.1000) ~ area + pop + pop18 + pop65 + docs + beds + poverty + pcincome + totalinc + region, data = crime_df_1)

boxcox(step.modelsqrt)
```

```{r}
plot(step.model, which = 4)

crimeOut = crime_df_1[-c(1,6),]

plot(crime_df_1$pcincome, crime_df_1$CRM.1000)
plot(crimeOut$pcincome, crimeOut$CRM.1000)

with = lm(sqrt(CRM.1000) ~ area + pop + pop18 + pop65 + docs + beds + 
    poverty + pcincome + totalinc + region, data = crime_df_1)

without = lm(sqrt(CRM.1000) ~ area + pop + pop18 + pop65 + docs + beds + 
    poverty + pcincome + totalinc + region, data = crimeOut)

plot(without)
```

```{r}
crime_df_10 = dplyr::select(crime_df_1, -bagrad, -hsgrad, -unemp)

corrplot(cor(crime_df_10), type = "upper", diag = FALSE)

pairs(crime_df_10)

check_collinearity(without)
```

Final model selected by the step-wise process

```{r}
final.step.model = lm(sqrt(CRM.1000) ~ area + pop + pop18 + pop65 + docs + beds + poverty + pcincome + totalinc + region, data = crimeOut)

summary(final.step.model)
```

Validation of stepwise before diagnostics and remedies
```{r}
set.seed(1)
train = trainControl(method = "cv", number = 5)

model_caret = train(CRM.1000 ~ area + pop + pop18 + pop65 + docs + beds + poverty + pcincome + totalinc + region, data = crime_df_1,
                   trControl = train,
                   method = 'lm',
                   na.action = na.pass)

model_caret$finalModel
print(model_caret)
```

```{r}
set.seed(1)
train = trainControl(method = "cv", number = 5)

model_caret = train(sqrt(CRM.1000) ~ area + pop + pop18 + pop65 + docs + beds + poverty + pcincome + totalinc + region, data = crimeOut,
                   trControl = train,
                   method = 'lm',
                   na.action = na.pass)

model_caret$finalModel
print(model_caret)
```

```{r}
confint(final.step.model)
confint(final.step.model,level = 0.90)
```